---
title: "Homework 3"
authour: Hunter Ross
date: 5/28/2025
format: html
---

https://github.com/Hunter-Ross-3/ENVS-193DS_homework-03.git



```{r -read in packages & data}
#| message: false
#loading packages
library(tidyverse)
library(janitor)
library(here)
library(flextable)
library(readxl)
library(png) 
library(grid) 
#loading data
my_data <- read.csv(here("data/my_data.csv"))
```

# Problem 1. Personal data

## a. Data summarizing

I calculated the Mean amount of money I spent on campus per day of the week. This summary allows me to compare how my spending varies depending on the day because on days that I'm on campus longer I expect to have higher spending. 

## b.Visualization

```{r}
#summarizing the mean spending by each day and naming it summary_data
summary_data <- my_data |> 
  group_by(day) |> 
  summarise(mean_spent = mean(spent, na.rm = TRUE))

#plotting data
ggplot(my_data, aes(x= day, y= spent)) +
 #making a jitter plot
   geom_jitter(width = 0.2, alpha = 0.6, color = "steelblue") + #adding specific width color and alpha to the jitterplot
  stat_summary(fun = mean, geom = "point", size = 4, color= "darkred") +#adding the mean into the jitter plot as the color red
#adding title of graph and axis
    labs(
    title = "Money Spent on Campus by Day",
    x = "Day of the Week",
    y = "Money Spent ($)") +
  # customizing the theme
  theme_minimal(base_size = 4) +
  theme(axis.title.x = element_text(size =14)) +
  theme(axis.text.x = element_text(size =10)) +
  theme(axis.title.y = element_text(size =14)) +
  theme(axis.text.y =  element_text(size =10)) +
  theme(title = element_text(size =14)) 
#changing the text size individually for all titles 
```

## c. Caption

Figure 1. Each point shows how much money I spent on campus each day. Red dots make the average for each day. I tend to spend more mid-week, especially on Thursday. 

## d. Table presentation

```{r}
#creating a flextable
summary_data |> #using the summay data
  rename('Day of Week' = day,#renaming day to day of the week 
         'Mean Spending($)'= mean_spent) |> #renaming mean_spent to mean spending 
  flextable() |> 
  autofit() |> 
  set_caption("Table 1. Average money spent on campus by weekday.") #creating a caption for the table
```


# Problem 2. Affective visualization

## a. Describe in words what an affective visualization could look like for your personal data

An affective visualization of this data could be a "spending spire," where each day is a ring in a circular calendar. The thickness of the ring would reflect how much i spent. It would also show how spending builds over time and could bring in the emotional data into my daily habits. 

## b. Sketch 


```{r}
grid.raster(readPNG(here("sketch.png")))
#the only way me and the DREAM lab could figure out how to get an image to actually show up
```


## c. Visualization draft

```{r}
# creating cicle graph
ggplot(my_data, aes(x = day, y = spent, fill = day)) + #using my original data not summary data
  geom_bar(stat = "identity", position = 
             "dodge", show.legend = FALSE) +
  #creating a base bar graph
  coord_polar() + #changing the bar graph to circle 
  theme_void() + #theme
  scale_fill_viridis_d() +
  labs(title = "Spending Spiral: Money Spent by Day")# title 
  theme(title = element_text(size =14))#changing the size of title

```


## d. Artistic statement
  -Content: This represents my daily spending as slices of a circle to visualize both emotional and routine-driven behavior. 
  -Influence: The circle graphs I’ve encountered in my professors' lectures and various articles have been a major source of inspiration for me. I find them to be an elegant and straightforward way to present information, combining clarity with visual appeal. 
  -Form: I plan to create my final circle graph in R using ggplot and using geom_bar with coord_polar to achieve the smooth circular visualization. I'll enhance the visual appeal by making sure to choose a nice color scheme, making clear labels, and ensuring an intuitive layout that interpreters the data both engaging and easy to 
  -Process: To make my ruff draft I started my calculating daily averages And then sketched the spiral. I used R coors_polar to test the visual structure and see the default colors. 

# Problem 3. Statistical critique

## a. Revisit and summarize

## b. Visual clarity

## c. Aesthetic clarity

## d. Recommendations







