---
title: "Homework 3"
format: html
---

https://github.com/Hunter-Ross-3/ENVS-193DS_homework-03.git



```{r -read in packages & data}
#| message: false
#loading packages
library(tidyverse)
library(janitor)
library(here)
library(gtable)
library(readxl)
#loading data
my_data <- read.csv(here("data/my_data.csv"))
```

## Problem 1. Personal data

# a. Data summarizing

I calculated the Mean amount of money I spent on campus per day of the week. This summary allows me to compare how my spending varies depending on the day because on days that I'm on campus longer I expect to have higher spending. 

# b.Visualization



```{r}
#summarizing the mean spending by each day

summary_data <- my_data |> 
  group_by(day) |> 
  summarise(mean_spent = mean(spent, na.rm = TRUE))

#ploting data
ggplot(my_data, aes(x= day, y= spent)) +
 #making a jitter plot
   geom_jitter(width = 0.2, alpha = 0.6, color = "steelblue") + 
  stat_summary(fun = mean, geom = "point", size = 4, color= "darkred") +
#adding title of graph and axis
    labs(
    title = "Money Spent on Campus by Day",
    x = "Day of the Week",
    y = "Money Spent ($)"
  ) +
  # customizing the theme
  theme_minimal(base_size = 4) +
  theme(axis.text.x= element_text(angle = 45, hjust = 1))
```

# c. Caption

# d. Table presentation

## Problem 2. Affective visualization

# a. Describe in words what an affective visualization could look like for your personal data

# b. Sketch (on paper)

# c. Visualization draft

# d. Artistic statement

## Problem 3. Statistical critique

# a. Revisit and summarize

# b. Visual clarity

# c. Aesthetic clarity

# d. Recommendations







